---
title: "07-4 Causal Inference"
author: "John Fieberg" 
---

We will consider a data set from an [**randomized experiment**](https://ecommons.cornell.edu/server/api/core/bitstreams/c739fdfa-a428-411b-9594-4acfbbc272ae/content) that evaluated the effect of server posture (standing vs. squatting) on the size of tip left by restaurant customers. The study was conducted by a server, who flipped a coin to randomly determine whether they would stand or squat when they first visited a table and introduced themselves to their customers. All subsequent interactions were performed from a standing position. The server also recorded additional information that might influence the size of the the tip. 

The data set is contained in the `experimentr` package. The code below will load the data and format some of the variables so that they are easier to understand:


```{r warning=FALSE, message= FALSE}
library(experimentr)
library(dplyr)
data(lynn)
tipping <- lynn %>% mutate(Shift = ifelse(daytime==0, "Day", "Evening"),
                        MaleFemale = ifelse(female== 1, "Female", "Male"),
                        Posture = ifelse(crouch==1, "Crouch", "Standing"),
                        Payment = ifelse(paid_by_credit_card==1, "Credit Card", "Cash")) %>%
  rename(Groupsize = groupsize, Tip = tip, Bill = bill)%>%
  select(Groupsize, Bill, Tip, Shift, MaleFemale, Posture, Payment)

```

Now, the `tipping` data set contains the following variables:

- `Groupsize` = number of customers dining at table
- `Bill` = total bill amount in dollars
- `Tip` = tip amount in dollars
- `Shift` = whether the customer or group of customers dined during a Day or Evening shift
- `MaleFemale` = Whether the bill was paid for by a Male or Female customer (note, other gender identifications were unfortunately not included in the experiment)
- `Posture` = whether the server was Crouched or Standing
- `Payment` = whether the bill was paid for using a Credit card or Cash. 

1. Create a DAG that represents possible connections between of the different variables in the data set.  Remember, these data come from a randomized experiment. That *should* influence the connections between `Posture`  and the other explanatory variables in the data set.

 
```{r }
library(ggdag)
library(ggplot2)
dag_object <- ggdag::dagify(tip~payment+bill+posture+gender,
                            payment~bill+groupsize, 
                            bill~groupsize+shift+posture,
                            groupsize~shift)
ggdag::ggdag(dag_object)
```
 


2. Use linear regression to estimate the **direct effect** of posture on the amount that customers tip.  

 

```{r}
 
```

3. Use linear regression to estimate the **total effect** of posture on the amount that customers tip.  (3 pts)

 
```{r}
 
```


## Compare to other model selection approaches

Let's look at stepwise backwards selection
```{r}
MASS::stepAIC(lm(Tip~Groupsize+Bill+Shift+MaleFemale+Posture+Payment, data=tipping))
```

Let's see if Augmented Backwards selection puts us somewhere differently. For this to work, it appears that we have to create our own dummy variables. This ends up in the same place...

```{r warning=FALSE, message=FALSE}
library(abe)
tipping2 <- tipping %>% mutate(Imale = ifelse(MaleFemale=="Male", 0, 1),
                               Ievening = ifelse(Shift=="Evening", 0,1),
                               IStand = ifelse(Posture=="Standing", 0, 1),
                               ICreditC = ifelse(Payment=="Credit Card", 0,1))
fullmod<-lm(Tip~Groupsize+Bill+Ievening+Imale + IStand + ICreditC, data=tipping2, 
            x = TRUE, y = TRUE)
abe(fullmod, criterion = "AIC", verbose=TRUE, exp.beta=FALSE, data=tipping)
```


We could also look at forward stepwise selection using:

```{r}
min.model <- lm(Tip ~ 1, data=tipping)
step(min.model, scope=( ~ Groupsize+Bill+Shift+MaleFemale+Posture+Payment), direction="forward", data=tipping)
```
This gets us to the same place.

# Bootstrap Stability

```{r warning=FALSE, message=FALSE}
library(rms)
set.seed(130)
fullmod.ols<-ols(Tip~Groupsize+Bill+Shift+MaleFemale+Posture+Payment, data=tipping,
                 x = TRUE, y = TRUE)  
validate(fullmod.ols, bw = TRUE) 
```


## Effect of groupsize

When using stepwise algorithms, groupsize is never chosen.  Use your DAG to explain why.
