---
title: "Rmd Example"
author: "Your name (or leave blank for HW)"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: pdf_document
---

To complete this exercise, you may need to install the following R packages:

- here 
- formatR

## Preamble

```{r setup}
# set global chunk options: images will be 7x5 inches
# wrap output at 60 characters 
library(knitr) # for creating reproducible reports
library(formatR) # for formating code in output file
library(here) # to facilitate reading in data using a relative path
opts_chunk$set(fig.width=7, fig.height=5, tidy.opts = list(width.cutoff = 60), tidy = TRUE)  
```

## Data

Enter code to read in data using a relative path

```{r}
coviddat <- read.csv(here("FW8051Data", "minnesota-history.csv"))
```



## Analysis

Plot the number of deaths over time. This will require that you grapple with a date-time variable in R. Some resources that may help if you struggle to figure out how to work with dates in R include:

- https://www.statmethods.net/input/dates.html
- https://r4ds.had.co.nz/dates-and-times.html
- your colleagues in this class!

```{r}

```


## Footer

We can create a reproducible using:
 
rmarkdown::render(here::here("Rcode", "filename"), output_dir = "Output", output_format="pdf_document")
  
It is always a good idea to include information about your R environment at the time you produced the report.

```{r}
sessionInfo()
```

